---
title: "README"
output: md_document
---

```{r global_options, echo=FALSE}                                               
knitr::opts_chunk$set(fig.path="figures/", warning=FALSE, message=FALSE, error=FALSE)
``` 

# growthcurve

growthcurve is an [R](http://r-project.org) package for analyzing biological growth curves.
It is designed to integrate into modern workflows, allowing it to be used in conjunction with other tools.

This package is currently a wrapper for the [grofit](http://cran.r-project.org/web/packages/grofit/index.html) package, which is no longer being developed.
This is temporary, as I plan to eventually make growthcurve an independent tool with more flexibility.

growthcurve is released with a [Contributor Code of Conduct](CONDUCT.md).      
By participating in this project, you agree to abide by its terms.


## Installation

`growthcurve` is not quite ready to be available on [CRAN](http://cran.r-project.org), but you can use [devtools](http://cran.r-project.org/web/packages/devtools/index.html) to install the current development version:

    if(!require("devtools")) install.packages("devtools")
    devtools::install_github("briandconnelly/growthcurve", build_vignettes=TRUE)

**Note that a lot of changes are being made, so things might not always work.**


## Quick Example

```{r quick_example, results="hide"}
library(growthcurve)
library(dplyr)
library(magrittr)

# Load the included 'pseudomonas' dataset
data(pseudomonas)

# Fit a logistic growth curve for PAO1 replicate 1
fit <- pseudomonas %>%
    filter(Strain == 'PAO1') %>%
    filter(Replicate == 1) %>%
    fit_growth_logistic(Time, CFUmL)

#plot(fit, show_raw=TRUE)
grofit::plot.gcFitModel(fit$grofit)
```

## Vignettes

The included vignettes contain more complete examples of how `growthcurve` can
be used:

* [Fitting a logistic growth curve](vignettes/logistic-growth.Rmd)
* Comparing growth


## Related Links
* [grofit](http://cran.r-project.org/web/packages/grofit/index.html)


## License

growthcurve is released under the Simplified BSD License.
