---
title: "README"
output: md_document
---

```{r global_options, echo=FALSE}                                               
knitr::opts_chunk$set(fig.path="figures/", warning=FALSE, message=FALSE, error=FALSE)
``` 

# growthcurve

[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)

growthcurve is an [R](http://r-project.org) package for analyzing [biological growth](https://en.wikipedia.org/wiki/Bacterial_growth), or "growth curves".
It is designed to integrate into modern workflows, allowing it to be used in conjunction with other tools.

This package is currently a wrapper for the [grofit](http://cran.r-project.org/web/packages/grofit/index.html) package, which is no longer being developed.
This is temporary, as I plan to eventually make growthcurve an independent tool with more flexibility.

growthcurve is released with a [Contributor Code of Conduct](CONDUCT.md).
By participating in this project, you agree to abide by its terms.


## Installation

growthcurve is not quite ready to be available on [CRAN](http://cran.r-project.org), but you can use [devtools](http://cran.r-project.org/web/packages/devtools/index.html) to install the current development version:

```{r intallation, eval=FALSE}
    if(!require("devtools")) install.packages("devtools")
    devtools::install_github("briandconnelly/growthcurve", build_vignettes=TRUE)
```
    
**Note that a lot of changes are being made, so things might not always work.**


## Fitting Growth Curves

growthcurve's most important function is `fit_growth`, which fits a growth curve to the given data.
Here, we'll fit a growth curve to one replicate population from the included `pseudomonas` data set, which has columns `Time` and `CFUmL`:

```{r fit_growth, message=FALSE}
library(dplyr)
library(growthcurve)

rep1 <- filter(pseudomonas, Replicate == 1 & Strain == "PAO1")
myfit <- fit_growth(rep1, Time, CFUmL)
```

Even better, we can do this all at once with pipes:

```{r fit_growth-pipe, message=FALSE}
myfit <- pseudomonas %>%
    filter(Replicate == 1 & Strain == "PAO1") %>%
    fit_growth(Time, CFUmL)
```


By default, `fit_growth` will try a few [parametric models](https://en.wikipedia.org/wiki/Parametric_model) and return the best one according to [AIC](https://en.wikipedia.org/wiki/Akaike_information_criterion).
The `type` argument can be used to use a specific model type.
Here, we'll use a logistic function:

```{r fit_growth-logistic, message=FALSE}
library(dplyr)
library(growthcurve)

rep1 <- filter(pseudomonas, Replicate == 1 & Strain == "PAO1")
myfit <- fit_growth(rep1, Time, CFUmL, type = "logistic")
```

Other options include `gompertz`, `gompertz.exp`, `richards`, and `spline`.
These can also be done directly using `fit_growth_gompertz`, `fit_gowth_gompertz.exp`, `fit_growth_richards`, and `fit_growth_spline`.
There's also a `fit_growth_parametric`, which finds the best among the parametric models.

## Interpreting Results

coming soon!


## Visualizing Growth Curves

growthcurve includes tools for visualizing growth curves using either R's base graphics or [ggplot2](https://cran.r-project.org/web/packages/ggplot2/index.html).

```{r base_example}
plot(myfit, show_raw=TRUE, show_maxrate=TRUE, show_asymptote=FALSE)
```

With ggplot, we can add growth curves to plots using `stat_growthcurve`.

```{r ggplot}
library(ggplot2)

pao1data <- filter(pseudomonas, Strain == "PAO1")
ggplot(data = pao1data, aes(x = Time, y = CFUmL, color = Replicate)) +
    geom_point(shape = 1) +
    stat_growthcurve()
```


## Vignettes

The included vignettes contain more complete examples of how `growthcurve` can
be used:

* [Fitting a logistic growth curve](vignettes/logistic-growth.Rmd)
* Comparing growth - coming soon!


## Related Links
* [grofit](http://cran.r-project.org/web/packages/grofit/index.html)


## License

growthcurve is released under the Simplified BSD License.
