---
title: "README"
output: md_document
---

```{r global_options, echo=FALSE}                                               
knitr::opts_chunk$set(fig.path="figures/", warning=FALSE, message=FALSE, error=FALSE)
``` 

# growthcurve

growthcurve is an [R](http://r-project.org) package for analyzing biological growth curves.
It is designed to integrate into modern workflows, allowing it to be used in conjunction with other tools.

This package is currently a wrapper for the [grofit](http://cran.r-project.org/web/packages/grofit/index.html) package, which is no longer being developed.
This is temporary, as I plan to eventually make growthcurve an independent tool with more flexibility.


## Installation

`growthcurve` is not quite ready to be available on [CRAN](http://cran.r-project.org), but you can use [devtools](http://cran.r-project.org/web/packages/devtools/index.html) to install the current development version:

    if(!require("devtools")) install.packages("devtools")
    devtools::install_github("briandconnelly/growthcurve")

**Note that a lot of changes are being made, so things might not always work.**


## Examples

### Fit a single logistic growth curve

Let's take `growthcurve` for a spin by fitting a logistic growth curve to some data from the included `phage` data set.

```{r sampledata1, results='hide'}
sampledata <- data.frame(Time=1:30, OD600=1/(1+exp(0.5*(15-1:30)))+rnorm(30)/20)
library(growthcurve)

# Load the data and take a look at the first few rows:
data(phage)
head(phage)
```

```{r sampledata1_table, echo=FALSE, results="asis"}
pander::pandoc.table(head(phage))
```

In this experiment, [Luis Zaman](http://luis.labfab.cc) measured the growth of two strains of *E. coli* both with and without phage. Replicate populations were grown in the wells of a microtiter plate, and absorbance was measured every three minutes.

Although we eventually want to compare the growth of each population, we're going to focus on the population in well **A12** for now. We can get this subset of the data using the `filter` function from [dplyr](http://cran.r-project.org/web/packages/dplyr/index.html).

```{r sampledata_filter}
library(dplyr)

a12data <- filter(phage, Well=="A12")
```

Now, let's fit a logistic growth curve for this population:

```{r example_logistic}
a12fit <- fit_growth_logistic(a12data, Time, OD420)
```

Note that we could have gotten fancy and combined the last two steps using [magrittr's](http://cran.r-project.org/web/packages/magrittr/index.html) lovely pipe operator:

```{r example_both}
a12fit <- phage %>%
    filter(Well=="A12") %>%
    fit_growth_logistic(Time, OD420)
```

We can now look at the results to learn about the fit:
```{r example_logistic_params}
a12fit$max_growth
a12fit$max_rate
a12fit$lag_length
a12fit$integral
```

For this fit, the maximum growth value is 0.980573926 AU (`a12fit$max_growth`), the maximum growth rate (the slope) is 2.792067e-05 AU/s (`a12fit$max_rate`), the lag phase ends approximately time -8326.2947 (`a12fit$lag_length`) (obviously not quite right), and the area under the growth curve is about 73088.97 (`a12fit$integral`).

We can also plot the results:

```{r plot_example_logistic}
#plot(a12fit)
grofit::plot.gcFitModel(a12fit$grofit)
```

If we'd like to take a look at how well the fitted curve matches the data, we can plot its residuals:

```{r resid_example_logistic}
plot(x=a12fit$grofit$raw.time, y=resid(a12fit), xlab="Time", ylab="Residuals")
abline(h=0)
```

So maybe this fit isn't quite right for these data. We can instead use another method.

To be continued...

## Related Links
* [grofit](http://cran.r-project.org/web/packages/grofit/index.html)


## License

growthcurve is released under the Simplified BSD License.
