---
title: "README"
output: md_document
---

```{r global_options, echo=FALSE}                                               
knitr::opts_chunk$set(fig.path='figures/', warning=FALSE, message=FALSE, error=FALSE)
``` 

# growthcurve

growthcurve is an [R](http://r-project.org) package for analyzing biological growth curves.
It is designed to integrate into modern workflows, allowing it to be used in conjunction with powerful tools like [dplyr](http://cran.r-project.org/web/packages/dplyr/index.html) and [magrittr](http://cran.r-project.org/web/packages/magrittr/index.html).

This package is currently a wrapper for the powerful [grofit](http://cran.r-project.org/web/packages/grofit/index.html) package, which is no longer being developed.
This is temporary, as I plan to eventually make growthcurve an independent tool with more flexibility.


## Installation

`growthcurve` is not quite ready to be available on [CRAN](http://cran.r-project.org), but you can use [devtools](http://cran.r-project.org/web/packages/devtools/index.html) to install the current development version:

    if(!require('devtools')) install.packages('devtools')
    devtools::install_github('briandconnelly/growthcurve')


## Examples

### Fit a logistic growth curve

First, let's create some sample data containing some time points in the *Time* column and some corresponding growth measurements in the *OD600* column and take a look at the first few rows:

```{r sampledata1}
sampledata <- data.frame(Time=1:30, OD600=1/(1+exp(0.5*(15-1:30)))+rnorm(30)/20)
```

```{r sampledata1_table, echo=FALSE, results='asis'}
pander::pandoc.table(head(sampledata))
```

(let's pretend like negative values for OD600 are ok for now)


Now, let's fit a logistic growth curve for this data set:

```{r example_logistic, results='hide'}
library(growthcurve)

lfit <- fit_growth_logistic(sampledata, Time, OD600)
```

Information about the logistic fit is available in `parameters`:

```{r example_logistic_params}
lfit$parameters
```

For this fit, the maximum growth value is 0.99848023 (`lfit$parameters$A`), the maximum growth rate (the slope) is 0.127109284 (`lfit$parameters$mu`), the lag phase ends approximately time 10.9985254 (`lfit$parameters$lambda`), and the area under the growth curve is about 15 (`lfit$parameters$integral`). The units for these results depend on the units in the input data.

We can also plot the results:

```{r plot_example_logistic}
plot(lfit)
```

If we'd like to take a look at how well the fitted curve matches the data, we can plot its residuals:

```{r resid_example_logistic}
lfit.res <- residuals(lfit)
plot(x=lfit$raw.time, y=lfit.res, xlab='Time', ylab='Residuals')
abline(h=0)
```


### Integrating other tools

As we saw earlier, our sample data contains some negative values. If we want to strip those out before fitting, we can first use `filter` from the excellent [dplyr](http://cran.r-project.org/web/packages/dplyr/index.html) package and then pipe the results to our fitting function using the pipe operator (`%>%`) from [magrittr](http://cran.r-project.org/web/packages/magrittr/index.html). While we're at it, maybe it takes a while for our machine to warm up, so let's also only use the data after time point 5.

```{r filtered, results='hide'}
library(dplyr)
library(magrittr)
library(growthcurve)

sampledata <- data.frame(Time=1:30, OD600=1/(1+exp(0.5*(15-1:30)))+rnorm(30)/20)
lfit2 <- sampledata %>%
    filter(OD600 >= 0) %>%
    filter(Time > 5) %>%
    fit_growth_logistic(Time, OD600)
    
plot(lfit2)

```



## Feature Requests and Bug Reports

For all feature requests and bug reports, visit [growthcurve on GitHub](https://github.com/briandconnelly/growthcurve/issues).  


## Related Links
* [grofit](http://cran.r-project.org/web/packages/grofit/index.html)


## License

growthcurve is released under the Simplified BSD License.
